---
title: "trophic_iso_ecol_workflow"
author: "Scott Yanco"
date: "April 22, 2019"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{checkyourself}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Questions about this vignette or the `chekyourself` package should be directed to Scott Yanco at: Scott.Yanco@ucdenver.edu  

# Introduction
This vignette demonstrates how the `checkyourself` package can be used to build and run a simple simulation model of consumer diet mixing.  Specifically, we simulate the process of trying to recover diet composition of a consumer via stable isotope analysis.  Diet end points' isotopic composition is "measured" from a smaple of varying sizes from each simulated potential prey population.  We explore the effect that estimation uncertaintly (with respect to diet end source isotopic composition) has on our ability to correctly parse the diet components from the consumer isotope mixture distribution.

#Set Up and Run Simulations

We first set up the set of declarations necessary to run the simulation model. In this construct we will simulate three potential prey populations of 10,000 individuals each. Each has a unique mean $\delta^{13}C$ and $\delta^{15}N$; all populations have equal isotopic variances. We supply a "true" proportion of total diet contributed by each potential source population which is the response variable we are ultmiately trying to correctly recover. We simulate a population of 25 consumers consuming prey 100 times each. Finally, we set a vector of sample sizes to draw from each potential prey source population to estimate the diet end point isotope compositions.  This set of samples will be the parameter space we explore with the model in this vignette.

```{r}
library(checkyourself)
num_sources <- 3 #number of diet endpoints to simulate
popsize <- c(10000,10000,10000) #size of each diet source pop to simulate
mu_carb <- c(-1, -10, -30) #vector of "true" mean carbon values for each source
sd_carb <- c(3,3,3) #vector of "true" carbon source standard deviations
mu_nit <- c(10,6,11) #vector of "true" mean nitrogen values for each source
sd_nit <-c(3,3,3) #vector of "true" nitrogen source standard deviations
diet_prop <- c(.1, .25, .65) #"true" proportional diet composition
steps <- 100 #number of steps per model run
num_individ <- 25 #number of consumers to simulate
#vector of sample sizes to use to estimate diet source isotope composition
sourcesamples <- c(3, 5, 10, 100) 
```

We can now simulate the "true" diet end source populations and then sample from each population (drawing samples of size `sourcesamples`).
```{r}
#make the food sources
food <- makefood(3, popsize, mu_carb, sd_carb, mu_nit, sd_nit)

# save diet source data to file for import to MixSIAR
#modify filepath to desired directory relative to workingdirectory.  
#Do not provide file extension in filename
samplesourcesvector(n_vec = sourcesamples, food = food, 
                    filepath = "../test_data/sources", writefile = T, 
                    returnobject = F)
```

`MixSIAR` reuires a file to establish discrimination factors.  For this simulation we consider no trophic discrimination (fixed at 0).  Below we create the necessary file.

```{r}
#create and save discrimination factor data
savediscrimination(food, filename="../test_data/simulated_discrimination.csv")
```

Now we can simulate the population of consumers "eating prey" and, subsequently, collect and save the measured isotope values from the simulated consumer tissue. The consumers stochastically consume from the prey populations with the probabilities supplied by `diet_prop`. The consumer isotope values reflect a weighted average of their consumed prey over the length of time defined by `window`, measured at the explicit time step defined by `time`. Here we set the collection time to the final step of the model and collect tissue samples with isitope values integrated over a 10-step window.

```{r}
#simulate agents over time
specimens <- eatfood(num_individ = num_individ, steps = steps, 
                     num_sources = num_sources, diet_prop = diet_prop, 
                     food = food)

#collect "observed" isotope values
obs_iso <- getiso(specimens, time = steps, window = 10)

#format to data frame ans save as csv for import to MixSIAR
formatiso(obs_iso, filename = "../test_data/simulated_iso.csv")
```

#Analysis

We employ standard trophic iso-ecology analysis tools (`MixSIAR`) to attempt to recover the consumer diet source proportions from the simulated data, given varying sample sizes for the estimation of diet source isotope distributions.

First, we load in the data, including separate objects for our multiple estimates of diet source isotope distributions
```{r}
library(MixSIAR)
library(splancs)
library(R2jags)

#load consumer/mixture data
mix <- load_mix_data(filename="../test_data/simulated_iso.csv",
                     iso_names=c("d13C","d15N"),
                     factors=NULL,
                     fac_random=NULL,
                     fac_nested=NULL,
                     cont_effects=NULL)

#load diet source data
source3 <- load_source_data(filename="../test_data/sources_3.csv",
                           source_factors=NULL,
                           conc_dep=FALSE,
                           data_type="means",
                           mix)
source5 <- load_source_data(filename="../test_data/sources_5.csv",
                           source_factors=NULL,
                           conc_dep=FALSE,
                           data_type="means",
                           mix)
source10 <- load_source_data(filename="../test_data/sources_10.csv",
                           source_factors=NULL,
                           conc_dep=FALSE,
                           data_type="means",
                           mix)
source100 <- load_source_data(filename="../test_data/sources_100.csv",
                           source_factors=NULL,
                           conc_dep=FALSE,
                           data_type="means",
                           mix)
#load discrimination data
discr <- load_discr_data(filename="../test_data/simulated_discrimination.csv", mix)
```

We now build a JAGS model for each diet source composition estimate and fit the model.

```{r}
# # Write the JAGS model file
# model_filename <- "../test_data/JAGS_model_1.txt"
# resid_err <- FALSE
# process_err <- TRUE
# write_JAGS_model(model_filename, resid_err, process_err, mix, source1)
# #run the model (using a 'test' lengthed chain)
# jags.1 <- run_model(run="very short", mix, source1, discr, model_filename,
#                     alpha.prior = 1, resid_err, process_err)

model_filename <- "../test_data/JAGS_model_3.txt"
resid_err <- FALSE
process_err <- TRUE
write_JAGS_model(model_filename, resid_err, process_err, mix, source3)
jags.3 <- run_model(run="very short", mix, source3, discr, model_filename,
                    alpha.prior = 1, resid_err, process_err)

model_filename <- "../test_data/JAGS_model_5.txt"
resid_err <- FALSE
process_err <- TRUE
write_JAGS_model(model_filename, resid_err, process_err, mix, source5)
jags.5 <- run_model(run="very short", mix, source5, discr, model_filename,
                    alpha.prior = 1, resid_err, process_err)

model_filename <- "../test_data/JAGS_model_10.txt"
resid_err <- FALSE
process_err <- TRUE
write_JAGS_model(model_filename, resid_err, process_err, mix, source10)
jags.10 <- run_model(run="very short", mix, source10, discr, model_filename,
                    alpha.prior = 1, resid_err, process_err)
```

```{r}
#set JAGS output options
output_options <- list(summary_save = TRUE,
                       summary_name = "summary_statistics",
                       sup_post = FALSE,
                       plot_post_save_pdf = TRUE,
                       plot_post_name = "posterior_density",
                       sup_pairs = FALSE,
                       plot_pairs_save_pdf = TRUE,
                       plot_pairs_name = "pairs_plot",
                       sup_xy = TRUE,
                       plot_xy_save_pdf = FALSE,
                       plot_xy_name = "xy_plot",
                       gelman = TRUE,
                       heidel = FALSE,
                       geweke = TRUE,
                       diag_save = TRUE,
                       diag_name = "diagnostics",
                       indiv_effect = FALSE,
                       plot_post_save_png = FALSE,
                       plot_pairs_save_png = FALSE,
                       plot_xy_save_png = FALSE)

#view the output
output_JAGS(jags.3, mix, source3, output_options)
output_JAGS(jags.5, mix, source5, output_options)
output_JAGS(jags.10, mix, source10, output_options)

params <- c("p.global")
summary(window(jags.3, start = 0))
jags.3$model



```

